---
title: "p8105_hw3_yl5508"
author: "Yifei LIU"
date: 2023/10/07
output: github_document
---

It's a proj for HW3 about VISUALIZATION AND EDA.

```{r setup, echo = FALSE, message = FALSE}
library(p8105.datasets)
library(tidyverse)
options(scipen = 999)
options(digits = 1)
```

## Problem 1

```{r load instacart}
#Variables review
data("instacart")
head(instacart, 5)
```

First, we shall take a quick look at the dataset.  
The `instacart` dataset contains `r nrow(instacart)` obs. and `r ncol(instacart)` variables.
For some key variables, the delineation for such is shown as below:  
- `order_id` is order identifier, `product_id` is product identifier, `user_id` is customer identifier. `product_name` is name of the product. `aisle_id` is aisle identifier. `aisle` is the name of the aisle. `department_id` is department identifier. `department` is the name of the department.   
- `reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise.  
- `order_number` is the order sequence number for this user (1=first, n=nth).  
- `order_dow` indicates the day of the week on which the order was placed. `order_hour_of_day` indicates the hour of the day on which the order was placed.  

```{r ins_clean and manipulate, message = FALSE}
#(a)
ins_clean =
  instacart |>
  janitor::clean_names()

#1
ais_num =
  ins_clean |>
  group_by(aisle_id, aisle) |>
  summarise(count = n()) |>
  arrange(desc(count))
head(ais_num, 5)
#the same as: 
#2
#ais_num =
  #ins_clean |>
  #group_by(aisle_id, aisle) |>
  #count(aisle, name = "count") |>
  #arrange(-count)
```

**(a)** `r ins_clean |> summarise(n_distinct(aisle_id)) |> pull()` (or using `r n_distinct(ins_clean |> pull(aisle))`) aisles are existed in data.  
`r head(ais_num, 1) |> pull(aisle)` is aisle that most items ordered from. The amount of items for this aisle is `r max(ais_num |> pull(count))`.  

```{r ins_ais_bargraph}
#(b)
ais_10k =
  ais_num |>
  filter(count > 10000) |>
  mutate(color = ceiling(count / 10000)) |>
  mutate(color = ifelse(color <5, as.character(color), "more"))

ais_10k_vertical =
  ais_10k |>
  ggplot(aes(x = reorder(aisle, -count), y = count, fill = factor(color), color = factor(color))) +
  #reorder ais_10k$aisle according to ais_10k$count. 
  #Actually, it is a factors reorder function, we can use parameter "FUN" to calculate the value in one factor as the reorder criteria.
  #'fill': color the inside, 'color': color the contour.
  geom_bar(stat = "identity", alpha = .75) +
  labs(
    title = "Number of Items Ordered in Each Aisle",
    x = "Aisle",
    y = "Number of Item",
  ) +
  viridis::scale_fill_viridis(discrete = TRUE, option = "viridis") +
  #connected with parameter 'fill'.
  viridis::scale_color_viridis(discrete = TRUE, option = "viridis") +
  #connected with parameter 'color'.
  #or using parameter 'scale_fill_hue(h = c(100, 400)) +'.
  theme_minimal() +
  theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5), axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
ais_10k_vertical

ais_10k_horizontal =
  ais_10k |>
  ggplot(aes(x = reorder(aisle, count), y = count, fill = factor(color), color = factor(color))) +
  geom_bar(stat = "identity", alpha = 0.75) +
  labs(
    title = "Number of Items Ordered in Each Aisle",
    x = "Aisle",
    y = "Number of Item",
  ) +
  viridis::scale_fill_viridis(discrete = TRUE, option = "viridis") +
  viridis::scale_color_viridis(discrete = TRUE, option = "viridis") +
  theme_minimal() +
  theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5), legend.position = "none") +
  coord_flip()
ais_10k_horizontal
```

```{r ggsave, message = FALSE}
ggsave("ais_10k_vertical.png", ais_10k_vertical)
ggsave("ais_10k_horizontal.png", ais_10k_horizontal)
```

**(b)** Two bar graphs are painted in this part. The data extends in different directions.  
For the generating process, we filtered the data so that they all lie in given range. Then we created factors for later coloring process. Use `ggplot` and `geom_bar` to build graph frameworks. After that, use function in `viridis` package to color the graphs.

```{r ins_most pop, message = FALSE}
#(c)
pop_num =
  ins_clean |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |>
  summarise(count = n()) |>
  arrange(aisle, -count) |>
  top_n(3, wt = count)
pop_num
```

**(c)** We filtered the data so that they all lie in given range. Then we group the data by variables `aisle` and `product_name`. Use `summarise` to generate the count of specific items given to the selected groups. At last, `top_n` is used to pick the 3 products with the biggest count in selected aisles.

```{r ins_ave h, message = FALSE}
#(d)
ave_h =
  ins_clean |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(ave_hour = mean(order_hour_of_day, na.rm = TRUE)) |>
  pivot_wider(names_from = order_dow, values_from = ave_hour)
ave_h
```

**(d)** We filtered the data so that they all lie in given range. Then we group the data by variables `product_name` and `order_dow`. Use `summarise` to generate the average hour of day given to the selected groups. At last, `pivot_width` is used to change the table shape, so it would meet the requirement mentioned in the problem.




